---
title: "STAT230 Work Team Project: Proposal & EDA"
author: "Work Team XX: NAMES"
date: "due Thursday, October 12th"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
---

```{r include=FALSE}
# Don't delete this chunk if you are using the mosaic package
# This loads the mosaic and mosaicData packages
library(mosaic)
library(car)
```

```{r include=FALSE}
# Some customization.  You can alter or delete as desired (if you know what you are doing).
# knitr settings to control how R chunks work.
require(knitr)
opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small"    # slightly smaller font for code
)
trellis.par.set(theme=theme.mosaic())  
options(digits=5)
```

[DELETE & DON'T INCLUDE the following instructions in your proposal]   

* Propose a proposal: The proposal part should include at least the following:  
    + Work Team number, all members’ names, and Work Team Project Title (if there is one).
    + Introduction & Motivation - What do you plan to study? What is the main (research) question of interest? WHY is this question interesting or important to your team?
    + Dataset - Where does it come from (including a citation of the data source)? What variables are you planning to use? Which is the response variable for your analysis?  
 
Since this is a warm-up project, I would recommend each work team to choose at least 4 variables but no more than 6 variables (including the response variable) from your dataset for the Work Team project.

* Perform an EDA (including graphs, numerical summaries, AND verbal descriptions) on the chosen variables that will be used in your modeling. Use appropriate graphical displays to justify that MLR is a reasonable choice for modeling (i.e. the CHOOSE step in 4-step modeling). You may find the document **Modeling in R: EDA** under *Useful Resources* particularly helpful for this deliverable.      
Additional Notes -   

This document should also include all the R code used to *wrangle* data, as well as those used to create the *EDA* output (plots, tables, numerical summaries, etc). Recall that you've learned various `R` functions for data wrangling back in **R Activity 2** (Part 3) and **R Activity 4** (Part 2), as well as from **R Tutorials** (under *Useful Resources*). Don't hesitate to ask for help from me or SDS fellows on this front.      

[--- DELETE until here: End of Instructions ---]


## Introduction & Motivation  


## Dataset & Wrangling   

On RStudio, be sure to save your dataset file in *the SAME folder* you saved this RMD file. I would strongly suggest that you create a specific folder on RStudio for this Project. When you wrangle your data, keep in mind that it's usually safer to save the mutated/filtered dataset as a *new* dataset.     

```{r}
#ds <- read.csv("XXX.csv")      ##read in your data
#dim(ds)                        ##size of the dataset
#names(ds)                      ##variable names   
```

```{r}
## wrangle!!  

```


## Exploratory Data Analysis (EDA)

Explore distributions and associations graphically and numerically.

